---
title: "AI Acceptance in the Sharing Economy"
subtitle: "Data Collection and Methodology --- Working Report"
author: "Htet Phone Khaing · Tatum Miranda Smith · Artem Cherkaskyy"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    number_sections: true
    latex_engine: xelatex
geometry: margin=2cm
fontsize: 11pt
header-includes:
  - \usepackage{float}
  - \usepackage{booktabs}
  - \floatplacement{figure}{H}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
suppressPackageStartupMessages({ library(dplyr); library(knitr) })

safe_read <- function(p) if (file.exists(p))
  tryCatch(read.csv(p, stringsAsFactors = FALSE), error = function(e) NULL) else NULL

fmt_range <- function(d) {
  d <- suppressWarnings(as.Date(d)); d <- d[!is.na(d)]
  if (!length(d)) return("n/a")
  paste(format(min(d), "%b %Y"), "--", format(max(d), "%b %Y"))
}
nice_platform <- function(x) {
  x <- sub("\\.com$", "", tolower(x))
  map <- c(turo="Turo", fatllama="Fat Llama", rvshare="RVshare", outdoorsy="Outdoorsy",
           lensrentals="Lensrentals", getaround="Getaround", kitsplit="KitSplit", sharegrid="ShareGrid")
  out <- map[x]; out[is.na(out)] <- x[is.na(out)]; unname(out)
}

v1 <- safe_read("data/reddit_all.csv"); raw <- safe_read("data/reddit_targeted.csv")
clean <- safe_read("data/reddit_clean.csv"); tp <- safe_read("data/trustpilot_reviews.csv")
v1_n <- if(!is.null(v1)) nrow(v1) else NA
raw_n <- if(!is.null(raw)) nrow(raw) else NA
clean_n <- if(!is.null(clean)) nrow(clean) else NA

if (!is.null(clean)) {
  cl_subs <- length(unique(clean$subreddit)); cl_threads <- length(unique(clean$url))
  cl_dates <- fmt_range(clean$date); cl_meanlen <- round(mean(nchar(clean$comment), na.rm=TRUE))
  cl_theme <- as.data.frame(table(clean$theme)); names(cl_theme) <- c("theme","n")
  top_subs <- clean |> count(subreddit, sort=TRUE) |> head(6)
  top_str <- paste(sprintf("r/%s (%d)", top_subs$subreddit, top_subs$n), collapse=", ")
}
if (!is.null(tp)) {
  tp_n <- nrow(tp); tp_dates <- fmt_range(tp$date_posted)
  tp_by_platform <- tp |> mutate(platform = nice_platform(platform)) |>
    group_by(Platform = platform) |>
    summarise(Reviews = n(), `Mean star` = round(mean(rating, na.rm=TRUE), 2),
              .groups="drop") |> arrange(desc(Reviews))
  tp_class <- table(cut(tp$rating, c(0,2,3,5), labels=c("neg","neu","pos")))
}
```

# The idea

Peer-to-peer rental marketplaces --- homes (Airbnb, Vrbo), cars and RVs (Turo,
Getaround, RVshare, Outdoorsy), and creative gear (Fat Llama, KitSplit,
Lensrentals) --- run entirely on **trust between strangers**, and they are now
inserting **AI and AI agents** into that relationship: automated guest
screening, chatbots, dynamic pricing, automated messaging. Our question:

> **How do users perceive AI agents in peer-to-peer rental platforms, and what
> drives or erodes their trust --- across rental types, and across open
> community discussion (Reddit) versus paying-customer reviews (Trustpilot)?**

We collect two complementary corpora. **Reddit is the baseline** --- where users
deliberate, complain and speculate in their own words. **Trustpilot is the
verification layer** --- structured reviews carrying a **1--5 star rating that
acts as a ground-truth sentiment label**. Where the open discussion and the
labelled reviews agree we gain confidence; where they diverge we learn something
about the gap between discourse and experience.

# Data source 1 --- Reddit (baseline)

Collected with `RedditExtractoR` (public Reddit API). The corpus went through
three passes, each fixing the previous one:

```{r reddit-passes}
passes <- data.frame(
  Pass = c("v1 keyword (global)", "v2 subreddit-targeted", "v2 cleaned + themed"),
  Script = c("scraping\\_data\\_v1.R", "01\\_scrape\\_reddit.R", "05\\_clean\\_reddit.R"),
  Rows = c(ifelse(is.na(v1_n),"~14,090",format(v1_n,big.mark=",")),
           ifelse(is.na(raw_n),"~1,009",format(raw_n,big.mark=",")),
           ifelse(is.na(clean_n),"~314",format(clean_n,big.mark=","))),
  Status = c("baseline only (off-topic)", "raw, on-topic", "USED for analysis"))
kable(passes, booktabs=TRUE, escape=FALSE,
      caption="Reddit passes. Files: reddit\\_all.csv, reddit\\_targeted.csv, reddit\\_clean.csv.")
```

**Why v1 failed.** A single global keyword search (`"AI rental platform helpful
automation trust"`) returned ~`r ifelse(is.na(v1_n),"14,000",format(v1_n,big.mark=","))`
mostly off-topic comments (K-pop, gaming, politics); under 1% mentioned
"rental". Kept only as a transparency baseline.

**How we targeted (v2).** We search *inside* topic-relevant subreddits instead
of all of Reddit, using **47 (subreddit, keyword) pairs** across six layers:

- **Home / stay** --- r/AirBnB, r/airbnb\_hosts, r/AirBnBHosts, r/vrbo, r/ShortTermRentals
- **Car / RV** --- r/Turo, r/gigwork, r/RVing, r/GoRVing
- **Camera / film / drone** --- r/photography, r/AskPhotography, r/cinematography, r/videography, r/Filmmakers, r/drones, r/dronephotography, r/DJI
- **Event tech** --- r/eventproduction, r/EventProduction, r/livesound
- **AI in commerce** --- r/Entrepreneur, r/SaaS, r/startups, r/automation, r/ChatGPT, r/OpenAI, r/artificial, r/singularity
- **Trust / privacy** --- r/privacy, r/technology

Keywords centred on AI, automation, hosting, rental and trust (e.g. "AI",
"host AI", "automated message", "rental AI", "AI agent"). Parameters: 30 threads
per search, `period = "all"`, threads kept only if `comments >= 6`, capped at
400 threads; deleted/removed and sub-20-character comments dropped. The keyword
that *finds* a thread is never used as its label --- sentiment is derived from
the text downstream.

**How we balanced and cleaned.** Two viral subreddits (r/Filmmakers,
r/Entrepreneur) hosted threads about AI replacing humans in creative/business
careers --- off-topic for rental trust --- so `05_clean_reddit.R` drops them and
tags each remaining comment `sharing_economy` or `ai_tech`. (The scrape can also
cap comments per thread at 120 to stop any single thread dominating.)

```{r reddit-desc, results='asis'}
if (!is.null(clean))
  cat(sprintf(paste0("**The dataset (`reddit_clean.csv`):** %s comments, %d ",
    "subreddits, %d threads, %s, mean length %d characters. Themes are close to ",
    "balanced (below); the top subreddits are %s.\n\n"),
    format(clean_n,big.mark=","), cl_subs, cl_threads, cl_dates, cl_meanlen, top_str))
```

```{r reddit-theme}
if (!is.null(clean)) {
  tt <- cl_theme; tt$Share <- paste0(round(100*tt$n/sum(tt$n)),"%")
  names(tt) <- c("Theme","Comments","Share")
  kable(tt, booktabs=TRUE, caption="Reddit theme balance.")
}
```

# Data source 2 --- Trustpilot (verification)

Trustpilot gives what Reddit cannot: a **1--5 star rating on every review** ---
a ground-truth label we did not have to infer. We can validate a sentiment
lexicon against these stars, then carry the validated lexicon to the unlabelled
Reddit comments (convergent validity).

**How we scraped it.** Trustpilot blocks naive scraping, so we used the Apify
actor casper11515/trustpilot-reviews-scraper via the **Apify web UI, once per
company**, downloading each result CSV into `data/trustpilot_raw/`.
`02_merge_trustpilot.R` then merges them (robust to BOM/UTF-8 export issues),
normalises columns, and keeps English reviews with real text and unique IDs ->
`data/trustpilot_reviews.csv`.

```{r tp-desc, results='asis'}
if (!is.null(tp))
  cat(sprintf(paste0("**The dataset:** %s unique English reviews across %d ",
    "platforms, spanning %s. Each row carries a star rating, title, body, ",
    "reviewer country and (where present) the company reply.\n\n"),
    format(tp_n,big.mark=","), nrow(tp_by_platform), tp_dates))
```

```{r tp-platform}
if (!is.null(tp))
  kable(tp_by_platform, booktabs=TRUE,
        caption="Trustpilot reviews per platform, with mean star rating.")
```

```{r tp-note, results='asis'}
if (!is.null(tp))
  cat(sprintf(paste0("Ratings collapse to **%d negative** (1-2 stars), ",
    "**%d neutral** (3), **%d positive** (4-5). The strong positive skew is ",
    "typical of opt-in review sites --- we down-sample the positive class for ",
    "balanced classification.\n\n"),
    tp_class[["neg"]], tp_class[["neu"]], tp_class[["pos"]]))
```

# Fit and next steps

The two sources triangulate: compute lexicon sentiment on Trustpilot text,
validate it against the star rating, apply the validated lexicon to Reddit, then
compare which *topics* drive negative sentiment in each. Agreement across two
structurally different corpora is stronger evidence than either alone.

**Planned analysis (owners to be assigned by the team):**

- **Preprocessing + topic modelling** --- clean, lemmatise, tokenise (1- and
  2-grams) both corpora; fit 3--4 LDA topics per theme (small corpus -> cap at
  four) plus a seeded LDA on trust words.
- **Visualisation + report** --- top-word bar charts, comparison/commonality
  word clouds, co-occurrence networks, a combined sentiment time-series; fold
  into the final paper.
- **Sentiment + embeddings + synthesis** --- AFINN/NRC sentiment validated
  against Trustpilot stars, NRC emotion profiles, GloVe neighbours of *trust* /
  *AI* / *host*, mapped to trust frameworks (Mayer 1995; TAM), ending in
  data-backed **AI-agent design cards**.

# Limitations

- Trustpilot skews positive (opt-in bias) --- we down-sample for classification.
- Equipment-rental platforms (Lensrentals, KitSplit) have few reviews ->
  weak per-platform claims there.
- One large r/ChatGPT thread pulls AI-side topics toward "AI tools"; few Reddit
  threads overall (38) cap LDA at 3--4 topics per theme.
- "Trust" is mostly *implicit* (host/renter, payment, review language), measured
  via topics and sentiment, not keyword counts. English-only.
- Reddit's anti-bot wall blocks fresh anonymous re-scrapes (needs an OAuth app;
  see `legacy/`). Turo/Getaround per-platform review dates need a parse check in
  `02_merge_trustpilot.R` before any per-platform time-series.

---

\footnotesize
Repository: <https://github.com/tyomachkaa/ai-acceptance-sharing-economy> ·
Online Content Analysis, WU Vienna, 2026.
