---
title: "AI Acceptance Across Service Platforms"
subtitle: "Data, Methodology and Progress against the Brief --- Working Report"
author: "Htet Phone Khaing · Tatum Miranda Smith · Artem Cherkaskyy"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    number_sections: true
    latex_engine: xelatex
geometry: margin=2cm
fontsize: 11pt
header-includes:
  - \usepackage{float}
  - \usepackage{booktabs}
  - \floatplacement{figure}{H}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
suppressPackageStartupMessages({ library(dplyr); library(knitr) })

safe_read <- function(p) if (file.exists(p))
  tryCatch(read.csv(p, stringsAsFactors = FALSE), error = function(e) NULL) else NULL
fmt_range <- function(d) {
  d <- suppressWarnings(as.Date(d)); d <- d[!is.na(d)]
  if (!length(d)) return("n/a")
  paste(format(min(d), "%b %Y"), "--", format(max(d), "%b %Y"))
}

rd <- safe_read("data/reddit_baseline.csv")        # Reddit discussion baseline
tp <- safe_read("data/trustpilot_flagged.csv")     # Trustpilot + ai_related flag
ai <- safe_read("data/ai_experience.csv")          # Trustpilot AI subset

if (!is.null(rd)) {
  rd_n <- nrow(rd); rd_subs <- length(unique(rd$subreddit))
  rd_dates <- fmt_range(rd$date); rd_len <- round(mean(nchar(rd$body), na.rm = TRUE))
  rd_cat <- rd |> count(platform_category, name = "Comments") |> arrange(desc(Comments))
  top_subs <- rd |> count(subreddit, sort = TRUE) |> head(6)
  top_str <- paste(sprintf("r/%s (%d)", top_subs$subreddit, top_subs$n), collapse = ", ")
}
if (!is.null(tp)) {
  tp_n <- nrow(tp); tp_dates <- fmt_range(tp$date_posted)
  tp_ai_n <- sum(tp$ai_related, na.rm = TRUE)
  tp_ai_star  <- round(mean(tp$rating[tp$ai_related], na.rm = TRUE), 2)
  tp_non_star <- round(mean(tp$rating[!tp$ai_related], na.rm = TRUE), 2)
}
```

# The idea

Online services increasingly put **AI and AI agents** between the user and the
thing they came for: chatbots answer support tickets, algorithms screen and
price, automated systems make decisions. We study **how people accept (or
reject) that AI** --- and, crucially, how acceptance changes with the *role* AI
plays. We compare three service contexts:

- **AI-native services** --- the product *is* an AI agent (ChatGPT, Claude,
  Replika, Gemini, Perplexity);
- **Peer-to-peer rentals** --- AI is a *peripheral* automation layer on a human
  marketplace (Turo, Airbnb);
- **Customer service** --- AI runs the *support* layer (chatbots replacing
  agents).

> **Research question.** How do users accept and experience AI agents across
> different service platforms --- and what drives or erodes trust as AI moves
> from being the product, to a marketplace add-on, to the support desk?

We use two complementary corpora. **Reddit is the discussion baseline** --- open
deliberation where users describe AI experiences in their own words.
**Trustpilot is the verification layer** --- reviews carrying a **1--5 star
rating** that acts as a ground-truth acceptance label. Where the open discussion
and the labelled reviews agree we gain confidence; where they diverge we learn
something real.

# Data source 1 --- Reddit (discussion baseline)

**The collection problem and fix.** Reddit's 2026 anti-bot wall blocks every
anonymous data endpoint (HTTP 403), which defeats the usual R/`RedditExtractoR`
and `.json`-based scrapers alike. We therefore pull from the **Arctic Shift**
research archive (a Pushshift successor) over its public API --- archived Reddit
data, no live blocking, no token or proxy (`01_fetch_reddit_arcticshift.R`).
AI-native subreddits are pulled in full; rental and customer-service subreddits
are pulled by AI/automation **keyword search** so we keep only on-topic comments.
Every comment is tagged with a `platform_category`.

```{r reddit-desc, results='asis'}
if (!is.null(rd))
  cat(sprintf(paste0("**The dataset (`reddit_baseline.csv`):** %s comments across ",
    "%d subreddits, %s, mean length %d characters. The top subreddits are %s. ",
    "Comments split across three service contexts:\n\n"),
    format(rd_n, big.mark = ","), rd_subs, rd_dates, rd_len, top_str))
```

```{r reddit-cat}
if (!is.null(rd)) {
  ct <- rd_cat; ct$Share <- paste0(round(100 * ct$Comments / sum(ct$Comments)), "%")
  names(ct)[1] <- "Platform category"
  kable(ct, booktabs = TRUE, caption = "Reddit comments by service context.")
}
```

A relevance check confirms the corpus is genuinely on-topic (the previous,
sharing-economy-targeted scrape was under 2%): customer-service comments are
~95% AI/automation-related and rental ~85% (keyword-filtered), while AI-native
comments are on-topic by subreddit context (people in r/ClaudeAI or r/replika
are discussing the AI product even when a comment omits the word "AI").

# Data source 2 --- Trustpilot (verification)

Trustpilot reviews carry a **1--5 star rating on every review** --- a
ground-truth acceptance label we did not have to infer. We scraped seven
peer-to-peer rental platforms via the Apify actor
`casper11515/trustpilot-reviews-scraper` (web UI, once per company),
merged them (`02_merge_trustpilot.R`), then flagged AI/automation reviews and
categorised platforms (`06_ai_acceptance.R`).

```{r tp-desc, results='asis'}
if (!is.null(tp))
  cat(sprintf(paste0("**The dataset:** %s reviews spanning %s. Of these, ",
    "**%d describe an AI / automation experience** (bots, algorithms, automated ",
    "decisions, \"no human\"). Those AI reviews average **%.2f stars vs %.2f** ",
    "for the rest --- on rental platforms, hitting automation correlates sharply ",
    "with frustration. This is the project's first concrete finding.\n\n"),
    format(tp_n, big.mark = ","), tp_dates, tp_ai_n, tp_ai_star, tp_non_star))
```

**Extending the verification layer.** The rental reviews give us AI-as-add-on.
To match the Reddit AI-native context we are scraping **AI-native platforms**
(character.ai, replika.com, openai.com, perplexity.ai) into the same pipeline;
`06_ai_acceptance.R` auto-tags them `ai_service`, giving Trustpilot the same
three-context structure as Reddit.

# Classes and how the sources fit

For the brief's per-class analysis our two **classes are positive vs negative
acceptance** --- on Trustpilot directly from the star rating (4--5 vs 1--2), on
Reddit from lexicon sentiment validated against those stars. **`platform_category`**
(AI-native / rental / customer-service) is the second lens applied to every
figure. The triangulation: compute sentiment on Trustpilot text, validate it
against the star label, apply the validated lexicon to Reddit, then compare
which topics drive negative sentiment in each context. Agreement across two
structurally different corpora is stronger evidence than either alone.

# Progress against the assignment brief

Each task is tagged **Done**, **Partly done**, **In progress** or **To do**.

**Dataset --- Done.** Two self-scraped corpora, no X/Twitter: Reddit
(`reddit_baseline.csv`, 4,250 comments, 3 categories) and Trustpilot
(3,912 star-labelled reviews + AI flag).

**Cleaning --- In progress.** Collection strips deleted/removed, short and
duplicate comments. Full normalisation (lowercase, punctuation/number/stop-word
removal, tokenisation, lemmatisation, 1- and 2-grams) is the next script
(`07_preprocess.R`).

**Word frequency & clouds --- To do.** Top-10 words per class as bar charts;
comparison and commonality clouds (uni- and bi-gram), per class.

**Word networks --- To do.** Per-class co-occurrence networks
(`widyr::pairwise_count` + `ggraph`).

**Sentiment --- To do.** AFINN/NRC on both corpora, both classes on one graph,
validated against Trustpilot stars (the convergent-validity check).

**Topic modelling (LDA) --- To do.** 3--5 topics per class, 1- and 2-gram, with
trust seed words and a `text2vec` cross-check.

**Embeddings --- To do.** GloVe / Word2Vec neighbours of *trust*, *AI*, *agent*,
*host* inside the topics.

**Synthesis & literature --- To do / partly identified.** Map findings to trust
and adoption frameworks (Mayer et al. 1995; Davis TAM; UTAUT), ending in
data-backed **AI-agent design cards**. Owners to be assigned by the team.

# Limitations

- Trustpilot skews positive (opt-in bias) --- we down-sample for classification.
- The rental keyword pull includes a broad `app` query that adds some general
  app-experience noise; every Reddit row keeps a `source_query` column so it can
  be filtered precisely.
- AI-native Trustpilot platforms are still being scraped, so the Trustpilot
  `ai_service` context is not yet populated.
- "Trust" is mostly *implicit* (bot/human, automated, algorithm language),
  measured via topics and sentiment, not keyword counts. English-only.
- Reddit data is archival (Arctic Shift); the live anti-bot wall prevents fresh
  anonymous scraping, and Turo/Getaround per-platform review dates need a parse
  check before any per-platform time-series.

---

\footnotesize
Repository: <https://github.com/tyomachkaa/ai-acceptance-sharing-economy> ·
Online Content Analysis, WU Vienna, 2026.
